A longitudinal analysis of the social information in infants' naturalistic

### 2022-01-07 paper/manuscript/manuscript_dec2021_apa_blinded.Rmd
incorporate last changes from google doc pass

We thus capitalize on this computational advance to annotate the entirety of this large, longitudinal video dataset. 
We use these detections to first report how the proportion of faces and hands changes with age and motor development milestones in each of the three children in the dataset. 
We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the head cameras as well as a diversity of locations in which videos were recorded.
Finally, making use of automated annotation of pose bounding boxes, we analyze the size, location, and variability of detected faces and poses across development in these three children.
Overall, these findings largely validate prior cross-sectional findings, and further suggest that automated annotated can be used to analyze naturalistic video datasets to build towards generalizable conclusions about the statistics of the social information in the infant view across a wide variety of populations.

@@

### Detector vs. human annotations of faces/hands
To test the validity of OpenPose's hand and face detections, we compared the accuracy detections relative to human annotations of 24,000 frames selected uniformly at random from videos of two children (S and A); this sample was conducted before child Y was added to the dataset.  Frames were jointly annotated for the presence of faces and hands by one author. A second set of coders recruited via AMT (Amazon Mechanical Turk) additionally annotated 3150 frames; agreement with the primary coder was >95%. Upon manually inspecting these 24K frames, we noticed that `r  num_frames_excluded` were sampled from videos taken from the allocentric perspective (i.e., not from the infant viewpoint); these frames and videos containing these frames were subsequently excluded from all other analyses.

@@

### Restricting detections to the center field-of-view
Given that the fish-eye lens may have distorted the edges of the visual field in the videos, we also examined whether we would see improved precision by restricting the included detections to the center field-of-view, and excluding detections that occured in the outer 20\% of each frame. This restricted field-of-view is also more comparable to video data from other head-mounted cameras. However, we found that excluding these detections did not improve the model's precision for either faces or hands; rather, it only decreased recall. Excluding peripehral detections led to an overall f-score of `r round(face_performance_center[3],2)`, for faces with a precision of `r round(face_performance_center[1],2)` and recall of `r round(face_performance_center[2],2)`. For hands, the F-score for center detections was `r round(hand_performance_center[3],2)`, with a precision of `r round(hand_performance_center[1],2)` and recall of `r round(hand_performance_center[2],2)`. Thus, detections that were made in the periphery of these videos were not more likely to be false alarms for either faces or hands. Appendix Figure A3 illustrates this finding, showing the density of false positives for both faces and hands that are concentrated in the center of the fied of view.

### Detector confidence
We next examined the degree to which including detections that the model was more or less confident in changed overall model performance for face/hand detection. When we restricted detections to high-confidence face/hand detections (>.5 confidence, default threshold for visualization in OpenPose) was not beneficial -- improving precision but dramatically impairing recall and thus overall performance: the F-score for high-confidence face detections was `r round(face_performance_hc[3],2)`, with a precision of `r round(face_performance_hc[1],2)` and recall of `r round(face_performance_hc[2],2)`; for high-confidence hand detections, the F-score was `r round(hand_performance_hc[3],2)`, with a precision of `r round(hand_performance_hc[1],2)` and recall of `r round(hand_performance_hc[2],2)`. Further analyses sweeping across the entire range of possible confidence detection thresholds revealed that including all detections was the most beneficial for optimizing the rate of true positives (see Appendix Figure A4 for full ROC curves for both face and hand detections).

@@

## Variability with motor development milestones
How does the social information in view change as a function of motor development? One idea is that children’s ability to stand and move on their own may fundamentally change the visual social information that they see (e.g., facial expressions, gaze cues, pointing)  [@walle2014infant; @walle2016infant] relative to infants who are still sitting and crawling.  Indeed, in in-lab studies, children's in-the-moment posture -- and that of their caregiver -- is a strong predictor of the visual social information that children see [@franchak2017see; @long2021automated].

We thus next analyzed how the social information in view changed as a function of each child reaching particular developmental motor milestones. Each parent in the SAYCam dataset completed the Ages and Stages Questionnaire (ASQ) [@bricker1999ages], which includes questions about their child's motor abilities, and we used the responses from this questionnaire to code the age (in months) at which each child became able to sit, cruise, and walk on their own (questions about crawling were not included in the questionnaire). We found considerable variability in the age at which each child achieved these three milestones. Participants S and A were already able to sit at 6 months of age, while Y began achieving this milestone at 8 months. A was able to cruise at 9 months of age, while S reached that milestone at 10 months, with Y just behind. Learning to walk showed the most variation, with S walking after 10 months, A after 13 months, and Y after 16 months of age.

We then used a discontinuity analysis to estimate whether the onset of each of these three motor milestones changed the proportion of faces/hands in each chid's view. Specifically, we used mixed-effects regressions with per-child random intercepts to estimate whether children's age (in months), each motor milestone, and their interaction significantly influenced the proportion of hands or faces that each child saw; see Appendix Tables A1-3 for full model specifications and tables with all coefficients. Overall, we found that the age at which these children began sitting did not significantly predict changes in either the amount of faces or hands that they saw. However, as children became able to cruise, they saw significantly more faces ($\beta = 0.27$) and hands ($\beta = 0.26$), suggesting that the ability to be upright and locomote even somewhat independently was sufficient to significantly change children's visual social environment. For faces, however, there was also a significant negative interaction of age and cruising ($\beta = -0.03$, $p=0.025$), and a marginal positive effect of age ($\beta = 0.03$, $p=0.055$). However, we did not find major differences with the onset of walking; as these children became able to walk, they saw marginally more hands ($\beta = 0.09$, $p=0.052$). Thus, these analyses confirm and extend cross-sectional findings, leveraging dense, within-child data to help elucidate the complex relationship between motor development and the social information in view.

@@

How might pose detection algorithms be improved upon such that they can achieve greater accuracy in analyzing infant egocentric videos?  This question is ripe for collaboration between computer vision experts and devleopmental psychologists. One promising direction may be to use detailed pose annotations of the people in the infant view to fine-tune later layers of pose detection networks. This has the potential to improve upon the overall accuracy of the network to detect the social information in infant egocentric videos. Standard computer vision models are rarely trained on the egocentric viewpoint, and we suspect that training these models on more naturalistic data may lead to more robust, generalizable detectors. A second potential area for innovation lies in leveraging the dense nature of these large-scale annotations to improve model accuracy: By pooling detections from surrounding frames, we may be able to build classifiers that perform more accurately on a single challenging viewpoint. Finally, a third area for improvement consists of building classifiers that can accurately distinguish between children's hands (as seen from the egocentric point of view) versus adults' hands--a notoriously challenging problem in computer vision [@bambach2015lending]. We expect that these goals may become more and more feasible as we also improve the head-mounted cameras that are used in both at-home and in-lab studies; increasing the resolution and quality of the images we record will no doubt improve the ability of computer vision algorithms to extract detailed information.

>>>>

We thus capitalize on this computational advance to annotate the entirety of this large, longitudinal video dataset. 
We use these detections to first report how the proportion of faces and hands changes with age in each of the three children in the dataset. 
We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the head cameras as well as a diversity of locations in which videos were recorded.
Finally, making use of automated annotation of pose bounding boxes, we analyze the size, location, and variability of detected faces and poses across development in these three children.
Overall, these findings largely validate prior cross-sectional findings, and further suggest that automated annotated can be used to analyze naturalistic video datasets to build towards generalizable conclusions about the statistics of the social information in the infant view across a wide variety of populations.

@@

### OpenPose vs. human annotations of faces/hands
To test the validity of OpenPose's hand and face detections, we compared the accuracy detections relative to human annotations of 24,000 frames selected uniformly at random from videos of two children (S and A); this sample was conducted before child Y was added to the dataset.  Frames were jointly annotated for the presence of faces and hands by one author. A second set of coders recruited via AMT (Amazon Mechanical Turk) additionally annotated 3150 frames; agreement with the primary coder was >95%. Upon manually inspecting these 24K frames, we noticed that `r  num_frames_excluded` were sampled from videos taken from the allocentric perspective (i.e., not from the infant viewpoint); these frames and videos containing these frames were subsequently excluded from all other analyses.

@@

### Restricting detections to the center field-of-view
Given that the fish-eye lens may have distorted the edges of the visual field in the videos, we also examined whether we would see improved precision by restricting the included detections to the center field-of-view, and excluding detections that occured in the outer 20\% of each frame. This restricted field-of-view is also more comparable to video data from other head-mounted cameras. However, we found that excluding these detections did not improve the model's precision for either faces or hands; rather, it only decreased recall. Excluding peripehral detections led to an overall f-score of `r round(face_performance_center[3],2)`, for faces with a precision of `r round(face_performance_center[1],2)` and recall of `r round(face_performance_center[2],2)`. For hands, the F-score for center detections was `r round(hand_performance_center[3],2)`, with a precision of `r round(hand_performance_center[1],2)` and recall of `r round(hand_performance_center[2],2)`. Thus, detections that were made in the periphery of these videos were not more likely to be false alarms for either faces or hands. Appendix Figure A3 illustrates this finding, showing the density of false positives for both faces and hands that are concentrated in the center of the field of view.

### Accuracy as a function of detection confidence
We next examined the degree to which including detections that the model was more or less confident in changed overall model performance for face/hand detection. When we restricted detections to high-confidence face/hand detections (>.5 confidence, default threshold for visualization in OpenPose) was not beneficial -- improving precision but dramatically impairing recall and thus overall performance: the F-score for high-confidence face detections was `r round(face_performance_hc[3],2)`, with a precision of `r round(face_performance_hc[1],2)` and recall of `r round(face_performance_hc[2],2)`; for high-confidence hand detections, the F-score was `r round(hand_performance_hc[3],2)`, with a precision of `r round(hand_performance_hc[1],2)` and recall of `r round(hand_performance_hc[2],2)`. Further analyses sweeping across the entire range of possible confidence detection thresholds revealed that including all detections was the most beneficial for optimizing the rate of true positives (see Appendix Figure A4 for full ROC curves for both face and hand detections).

@@

@@

How might pose detection algorithms be improved upon such that they can achieve greater accuracy in analyzing infant egocentric videos?  This step is ripe for collaboration between computer vision experts and developmental psychologists. One promising direction may be to use detailed pose annotations of the people in the infant view to fine-tune later layers of pose detection networks. This has the potential to improve upon the overall accuracy of the network to detect the social information in infant egocentric videos. Standard computer vision models are rarely trained on the egocentric viewpoint, and we suspect that training these models on more naturalistic data may lead to more robust, generalizable detectors. A second potential area for innovation lies in leveraging the dense nature of these large-scale annotations to improve model accuracy: By pooling detections from surrounding frames, we may be able to build classifiers that perform more accurately on a single challenging viewpoint. Finally, a third area for improvement consists of building classifiers that can accurately distinguish between children's hands (as seen from the egocentric point of view) versus adults' hands -- a notoriously challenging problem in computer vision [@bambach2015lending]. We expect that these goals may become more and more feasible as we also improve the head-mounted cameras that are used in both at-home and in-lab studies; increasing the resolution and quality of the images we record will no doubt improve the ability of computer vision algorithms to extract detailed information.
